Nonlinear Intrinsic Variables and State Reconstruction in Multiscale Simulations

### 2013-12-27 paper/revision3.tex
Added files related to paper

\section{Introduction}
The last decade has witnessed extensive advances in dimensionality reduction techniques:
finding meaningful low-dimensional descriptions of high-dimensional data \cite{Tenenbaum2000,Roweis2000,Donoho2003,Belkin2003,Coifman2006}.
%
Often, high-dimensional data do not uniformly fill the entire ambient space, but rather lie on a lower-dimensional manifold.
%
Discovering a parameterization of the manifold therefore yields a compact representation of the data, and may lead to efficient analysis and processing schemes.
%
In particular, these developments have the potential to significantly enable the computational exploration
of physicochemical problems.
%
Traditionally, state reduction in chemical reaction networks has been done using some {\em a priori} knowledge about separation of time scales,
for example, using the quasi-steady state approximation to reduce a network of chemical reactions\cite{bowen1963singular}.
%
Alternatively, empirical knowledge about the system of interest has been used to develop coarse grained descriptions,
such as grouping atoms of a macromolecule into larger ``beads'' to accelerate simulations\cite{monticelli2008martini, spiga2013electrostatic,izvekov2005systematic, saunders2013coarse}.
%
The use of manifold learning allows for the systematic discovery of data-driven, low-dimensional descriptions that circumvent the need for such {\em a priori} knowledge of the system.
%
If the (high-dimensional) data $\mathbf{Y}(t)$ arise from, for example, a
molecular dynamics simulation of a macromolecule in solution, or from the stochastic
simulation of a complex chemical reaction scheme, the detection of a few good, coarse-grained
``reduction coordinates" $\mathbf{x}(t)$ can be invaluable in understanding and predicting system behavior.
%
For example, such reduction coordinates may be the dihedral angles inferred from trajectories of the atoms of a macromolecule.

@@

Let $\mathbf{D}$ be a diagonal matrix whose elements are the row sums of $\mathbf{W}$, and let
\begin{equation}\label{eq:kernel}
\mathbf{W}^{\mathrm{norm}} = \mathbf{D}^{-1/2}\mathbf{W}\mathbf{D}^{-1/2}
\end{equation}
be a normalized kernel that shares its eigenvectors with the normalized graph-Laplacian $\mathbf{I}-\mathbf{W}^{\mathrm{norm}} \:$ \cite{chung1997spectral}.
%
The diagonal entries of $D$ serve as estimates for the density of data around each point.
%
The eigenvectors of $\mathbf{W}^{\mathrm{norm}}$, denoted $\psi_j$, reveal the underlying structure of the data \cite{coifman2005geometric}.
%
Specifically, the $i$-th coordinate of the $j$-th eigenvector can be associated with an intrinsic coordinate $j$ of the sample $\mathbf{x}(i)$ of the underlying process.
%
The eigenvectors are ordered such that $|\lambda_1| \ge |\lambda_2| \ge \dots \ge |\lambda_n|$, where $\lambda_j$ is the eigenvalue associated with eigenvector $\psi_j$.
%
Because $\mathbf{W}^{\mathrm{norm}} \sim \mathbf{D}^{-1}\mathbf{W} $, and $\mathbf{D}^{-1}\mathbf{W}$ is row-stochastic,
$\lambda_1 = 1$ and $\psi_1$ is the diagonal of $\mathbf{D}^{1/2}$.
%
The next few eigenvectors can be argued to describe the geometry of the underlying manifold \cite{coifman2005geometric}.
%
However, some eigenvectors can be higher harmonics of the same principal direction along the data manifold.
This is analogous to how the eigenfunctions $\cos x$ and $\cos 2x$ of the usual Laplacian in one spatial dimension
and with no flux boundary conditions are one-to-one with the values of $x$ for $0 \le x \le 1$;
one must check for correlations between the eigenvectors before selecting those that describe the underlying manifold geometry.
%
The above steps to construct the nonlinear intrinsic variables are summarized in Algorithm \ref{algo}.

@@

\subsection{Limiting Diffusion Operator} \label{sec:diff_limit}
Assume that the dynamics of the underlying variables in \eqref{eq:dynamics} can be rewritten as
\begin{equation} \label{eq:Langevin}
	d\mathbf{x}(t) = - \nabla U(\mathbf{x}(t)) dt + d \mathbf{w}(t),
\end{equation}
where the drift term is expressed as a gradient of a potential $U$ and $\dot{\mathbf{w}}(t) \in \mathbb{R}^d$ is a vector consisting of independent white noises.
%
Nadler {\em et al.} \cite{nadler2006diffusion} showed that in the limit $\varepsilon \rightarrow 0$ and $T \rightarrow \infty$ the kernel $\mathbf{W}^{\mathrm{norm}}$ converges to the continuous backward Fokker-Planck operator
\begin{equation}
	\mathcal{L} = \Delta - \nabla U \cdot \nabla.
	\label{eq:limiting_operator}
\end{equation}
%
We remark that, while in diffusion maps \cite{coifman2005geometric}, the limiting operator is the Laplacian on the manifold on which the observations lie,
here the Laplacian operator on the data that is based on the Mahalanobis distance converges to a limiting Laplacian operator on the
manifold of the intrinsic variables $\mathbf{x}$.

The convergence to the limiting operator implies that the eigenvectors of $\mathbf{W}^{\mathrm{norm}}$, which are used to represent the NIV, approximate the eigenfunctions of the limiting Fokker-Planck operator on the NIV manifold.
%
As such, according to \eqref{eq:limiting_operator}, they depend on the potential.
%
Thus, the constructed embedding is consistent as long as the subsets of observations $\mathbf{Y}(t)$ are
adequate samples of the same (equilibrium) density.
%
For example, when looking at multiple data sets from a molecular simulation, the embeddings will be consistent provided each simulation sufficiently samples the same underling free energy surface.

In order to merge observations from different regions,
the fundamental question is, to what extent the respective sampling densities affect the coordinates we obtain.
%
In principle, we can decouple diffusion dynamics (and thus, equilibrium densities)
from geometry by normalizing the diffusion kernel with the sample density.
%
This has been explored for diffusion maps independently of NIV \cite{coifman2005geometric}.
%
When the normalized kernel is defined as
\begin{equation} \label{eq:norm_kernel}
\widetilde{\mathbf{W}}^{\mathrm{norm}} = \mathbf{D}^{-1}\mathbf{W}\mathbf{D}^{-1},
\end{equation}
the limiting diffusion operator is the Laplace-Beltrami operator \cite{nadler2006diffusion}, given by $\widetilde{\mathcal{L}} = \Delta$.
%
In other words, this diffusion captures only the geometry of the observations and is invariant to the potential (or, alternatively,
to the details of the sampling density).
This kernel is similar to the kernel in \eqref{eq:kernel}, but $D$ appears with coefficient -1 rather than $-1/2$.

Since local coordinates are based on the noise term in \eqref{eq:Langevin}, sets of measurements from different regions
will essentially detect the same system of coordinates (the same variable ``axes");
yet different sampling densities (arising, for example,
from the drift term in \eqref{eq:Langevin}) will result in different {\em scalings} of these axes.
%
Applying the Laplace-Beltrami normalization in the NIV context will thus eliminate sampling density effects.
%
An alternative approach would find the scaling transformations that optimally ``register"
data sampled from common regions (in the spirit of histogram reweighting \cite{ferrenberg1988new}).

@@

A na\"{\i}ve way to extend $\mathbf{f}$ to a new data point $\Psi(\tilde{i})$ is find the point's nearest neighbors in NIV space and average their function values.
%
A different, point-wise adaptive approach is described by Buchman {\em et al.} \cite{buchman2011high}:
high-dimensional hurricane tracks were estimated from low dimensional embedding coordinates using a weighted average of the points close to $\Psi(\tilde{i})$ in the embedded space.
%
However, this point-wise adaptation requires setting the nearest neighborhood radius parameter for every point.
%
The LP algorithm finds the appropriate nearest neighborhood radius for each new point $\Psi(\tilde{i})$.
%
This radius will be large in smooth regions of the function, and small in regions in which $\mathbf{f}$ contains higher frequency components.
%
The LP approximation of a new, high dimensional point $\mathbf{Y}(\tilde{i})$ is calculated by a weighted average of the function values that belong to the neighboring points.
%
The weights are based on the pairwise distances in the intrinsic, low-dimensional space.
%
In practice, a set of smoothing operators $P_0, P_1, \ldots, P_l$, with
\begin{equation} \label{eq:LP_multi_scale_app}
P_l \propto e^{-d^2(\Psi(i),\Psi(j)) / 2^{-l} \sigma_0},
\end{equation}
are constructed and later extended to create the multiscale approximations as defined in \eqref{eq:LP_multi_scale}.
%
The LP algorithm for the inverse mapping is summarized in Algorithm \ref{algo_LP}.

@@

\item
Use the smoothing operators to obtain a multiscale representation (see \eqref{eq:LP_multi_scale_app}) of $\mathbf{f}:\Psi(i) \rightarrow Y(i)$.

@@

@@

We generate 3000 initial conditions $\mathbf{Y}_0(1), \dots, \mathbf{Y}_0(3000) \in \mathbb{R}^6$ uniformly at random from the
region of state space where all concentrations (respecting conservation laws) are non-negative.
%
We evolve each point $\mathbf{Y}_0(t)$ forward for 10 time units using the SSA to obtain a point $\mathbf{Y}(t) \in \mathbb{R}^6$;
according to the time scales calculated from the linearized ODEs, 10 time units is sufficiently long for the initial points in the ODE system to converge to the two-dimensional manifold,
but not long enough for the points to converge to a one-dimensional curve or to the final steady state (see Appendix \ref{app:rxn} for more details).
%
In our stochastic simulations, the initial points appear to converge to an approximate two-dimensional manifold
(in expected value, see Figure \ref{fig:rxn_manifolds}).
%
We consider $\mathcal{Y} = \{ \mathbf{Y}(t): t=1, \dots, 3000 \}$ to be representative points ``on" this apparent two-dimensional manifold.
%
From each manifold point $\mathbf{y} \in \mathcal{Y}$, we run 20 short simulation ``bursts'', each for 0.2 time units.
%
We denote the endpoints from the short simulations as $\mathcal{Y}^{burst}(\mathbf{y})$.

@@

To determine whether a sufficient amount of data is available to obtain an accurate embedding, we compute the NIV embeddings for different numbers of data points
(arising from different numbers of trajectories initialized randomly in the full concentration space).
%
Figure \ref{fig:rxn_convergence} shows the two-dimensional NIV embeddings and corresponding eigenvalue spectra of the kernel for the Gillespie SSA example computed from 10, 100, 500, and 1500 data points.
%
We observe that the eigenvalue spectra appear converged for 100 data points and above.
%
This implies that a sufficient amount of data is available merely to discover the intrinsic variables.
%
However, we observe that 100 points are insufficient to construct a self-consistent embedding {\em in the two variables discovered}.
%
In order to account for the different scaling of the NIV coordinate axes, we require at least 500 data points.
%
This last convergence is exemplified in Figure \ref{fig:rxn_convergence} (c) and (d).
%
The convergence of the spectrum of the kernel and the ultimate self-consistency of the NIV embedding is our empirical indicator that a sufficient amount of data is available.
\begin{figure*}[ht]
    \includegraphics[width=6in]{fig4b}
    \caption{Two-dimensional NIV embeddings (top) and corresponding eigenvalue spectra (bottom) calculated from subsets of $\mathcal{Y}_1$ containing (a) 10 data points, (b) 100 data points, (c) 500 data points, and (d) 1500 data points. One can see that the embedding and spectrum in (a) is inconsistent with the other embeddings, and that, although the spectrum in (b) appears converged, the embedding is not converged until (c). }
    \label{fig:rxn_convergence}
\end{figure*}

@@

We first compare NIV with direct diffusion maps \cite{coifman2005geometric}, an established nonlinear dimensionality reduction technique.
%
We consider 10,000 data points from our simulation $\mathbf{Y}(1), \dots, \mathbf{Y}(10000) \in \mathbb{R}^{30}$; every 100 data points comes from a continuous simulation trajectory.
%
We construct two data sets:
%
$\mathcal{Y}_{even} = \{\mathbf{Y}(t) \text{ restricted to atoms 2, 4, 6, 8, and 10}: t=1, \dots, 10000 \}$,
and $\mathcal{Y}_{odd} = \{\mathbf{Y}(t) \text{ restricted to atoms 5, 7, and 9}: t=1, \dots, 10000 \}$ (see Figure \ref{fig:ala_fes} for the atom indexing).
%
We then compute the NIV and diffusion maps embeddings for $\mathcal{Y}_{even}$ and $\mathcal{Y}_{odd}$;
for NIV, we compute the covariances as in \eqref{eq:cov}, with $L=10$.

@@

We then use NIV together with LP to predict the conformation of Ala2 when we only observe some of the atoms.
%
We have 20000 data points $\mathbf{Y}(1), \dots, \mathbf{Y}(20000)$, where every 100 data points come from one continuous simulation trajectory.
%
Our first data set (which will serve as our training data for LP), $\mathcal{Y}_{all}$,
consists of the first 10000 data points ($\mathcal{Y}_{all} = \{\mathbf{Y}(t): t=1, \dots, 10000\}$).
%
Our second data set (which will serve as our test data), $\mathcal{Y}_{odd}$, consists of the last 12000 data points restricted to {\em only} the odd atoms
($\mathcal{Y}_{odd} = \{ \mathbf{Y}(t) \text{ restricted to the odd atoms}: t = 8001, \dots, 20000\}$).
%
We compute the covariances as in \eqref{eq:cov} with $L=15$.
%
We compute the NIV embedding for the training data $\mathcal{Y}_{all}$ and the test data $\mathcal{Y}_{odd}$; we then use LP interpolation from the training data to predict the location of all the atoms for each point in the test data.

@@

@@

Figure \ref{fig:ala_recon} shows the reconstructed position from partial observation versus true position for certain selected atoms.
%
The strong correlation between the true and reconstructed positions is easier to appreciate for
atoms that move substantially within the data set (such as atoms 1 and 3).
%
Figure \ref{fig:ala_molecules} shows molecular structures for the true and reconstructed configurations for selected data points;
there is qualitative agreement between the true and reconstructed configurations.
%
The discrepancies between the true and reconstructed positions of atoms 9 and 10 (see Figure \ref{fig:ala_fes} for atom indexing) for some of the molecular structures are not surprising.
%
From Figure \ref{fig:ala_embed}, one can see that the first three NIV describe the flipping of atoms 1 and 3 and the dihedral angles $\phi$ and $\psi$.
%
However, atoms 9 and 10 do not participate in any of these physical quantities, and so there is little information about their positions contained in the first three NIV.
%
We suspect that using more NIV would result in a better description of atoms 9 and 10 and lead to more accurate structure reconstructions.

For a brief comparison of LP over other reconstruction techniques, we also reconstruct configurations from the
NIV components using simple nearest-neighbor interpolation.
%
The average reconstruction error, scaled by the average bond length within the molecule, is shown in Figure \ref{fig:ala_mse};
LP arguably outperforms simple nearest neighbor search for all of the atoms.
%
To determine why LP outperforms nearest neighbor search, we examine the errors at different scales in the LP algorithm.
%
The example described in Figure \ref{fig:LP_ex} demonstrates the appropriateness of the LP algorithm to signals with an intrinsic multiscale structure.
%
We perform a similar analysis using the data from the alanine dipeptide example, and show that the observed signal does exhibit a similar error structure,
with large errors in some regions of NIV space and small errors in other regions;
these results are shown in Figure \ref{fig:LapPyr_ala2_errors}.
%
We therefore conclude that the alanine dipeptide data does contain a measure of multiscale behavior, and, therefore, that LP is an appropriate algorithm for reconstruction.

@@

\begin{figure*}[ht]
    \centering
    \includegraphics[width=3in]{fig10b}
    \caption{The error in the reconstructed position of the z-coordinate of atom 10 as a function of $\psi_1$. One can see that the error is large when $\psi_1$ is large (similar to the error structure in Figure \ref{fig:LP_ex}), demonstrating that our data is indeed multiscale and that Laplacian Pyramids is an appropriate interpolation algorithm to use.}
    \label{fig:LapPyr_ala2_errors}
\end{figure*}

@@

Other issues, such as accurately estimating the covariance matrices required for the computation of the Mahalanobis distance, are also of current research interest.
%
It is clearly necessary to link this type of calculation with modern estimation techniques for (multiscale) diffusions \cite{ait2002maximum, ait2003effects, ait2008closed}
to test the appropriateness of the window sampling lengths selected; this will determine the accuracy of the noise covariance estimation by eliminating
the bias due to drift variations.
%
We are confident that the exploration of these open questions will enable the use of our methodology in many  interesting applications,
such as merging data from molecular simulations at different levels of granularity, or merging simulation data with experimental observations.

@@

%\bibliography{../../references/references}
%\bibliography{references}
%merlin.mbs aipnum4-1.bst 2010-07-25 4.21a (PWD, AO, DPC) hacked
%Control: key (0)
%Control: author (8) initials jnrlst
%Control: editor formatted (1) identically to author
%Control: production of article title (-1) disabled
%Control: page (0) single
%Control: year (1) truncated
%Control: production of eprint (0) enabled
\begin{thebibliography}{46}%
\makeatletter
\providecommand \@ifxundefined [1]{%
 \@ifx{#1\undefined}
}%
\providecommand \@ifnum [1]{%
 \ifnum #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \@ifx [1]{%
 \ifx #1\expandafter \@firstoftwo
 \else \expandafter \@secondoftwo
 \fi
}%
\providecommand \natexlab [1]{#1}%
\providecommand \enquote  [1]{``#1''}%
\providecommand \bibnamefont  [1]{#1}%
\providecommand \bibfnamefont [1]{#1}%
\providecommand \citenamefont [1]{#1}%
\providecommand \href@noop [0]{\@secondoftwo}%
\providecommand \href [0]{\begingroup \@sanitize@url \@href}%
\providecommand \@href[1]{\@@startlink{#1}\@@href}%

>>>>

\section{Introduction}
The last decade has witnessed extensive advances in dimensionality reduction techniques:
finding meaningful low-dimensional descriptions of high-dimensional data \cite{Tenenbaum2000,Roweis2000,Donoho2003,Belkin2003,Coifman2006}.
%
Often, high-dimensional data do not uniformly fill the entire ambient space, but rather lie on a lower-dimensional manifold.
%
Discovering a parameterizations of the manifold therefore yields a compact representation of the data, and may lead to efficient analysis and processing schemes.
%
In particular, these developments have the potential to significantly enable the computational exploration
of physicochemical problems.
%
Traditionally, state reduction in chemical reaction networks has been done using some {\em a priori} knowledge about separation of time scales,
for example, using the quasi-steady state approximation to reduce a network of chemical reactions\cite{bowen1963singular}.
%
Alternatively, empirical knowledge about the system of interest has been used to develop coarse grained descriptions,
such as grouping atoms of a macromolecule into larger ``beads'' to accelerate simulations\cite{monticelli2008martini, spiga2013electrostatic,izvekov2005systematic, saunders2013coarse}.
%
The use of manifold learning allows for the systematic discovery of data-driven, low-dimensional descriptions that circumvent the need for such {\em a priori} knowledge of the system.
%
If the (high-dimensional) data $\mathbf{Y}(t)$ arise from, for example, a
molecular dynamics simulation of a macromolecule in solution, or from the stochastic
simulation of a complex chemical reaction scheme, the detection of a few good, coarse-grained
``reduction coordinates" $\mathbf{x}(t)$ can be invaluable in understanding and predicting system behavior.
%
For example, such reduction coordinates may be the dihedral angles inferred from trajectories of the atoms of a macromolecule.

@@

Let $\mathbf{D}$ be a diagonal matrix whose elements are the row sums of $\mathbf{W}$, and let $\mathbf{W}^{\mathrm{norm}} = \mathbf{D}^{-1/2}\mathbf{W}\mathbf{D}^{-1/2}$
be a normalized kernel that shares its eigenvectors with the normalized graph-Laplacian $\mathbf{I}-\mathbf{W}^{\mathrm{norm}} \:$ \cite{chung1997spectral}.
%
The eigenvectors of $\mathbf{W}^{\mathrm{norm}}$, denoted $\psi_j$, reveal the underlying structure of the data \cite{coifman2005geometric}.
%
Specifically, the $i$-th coordinate of the $j$-th eigenvector can be associated with an intrinsic coordinate $j$ of the sample $\mathbf{x}(i)$ of the underlying process.
%
The eigenvectors are ordered such that $|\lambda_1| \ge |\lambda_2| \ge \dots \ge |\lambda_n|$, where $\lambda_j$ is the eigenvalue associated with eigenvector $\psi_j$.
%
Because $\mathbf{W}^{\mathrm{norm}} \sim \mathbf{D}^{-1}\mathbf{W} $, and $\mathbf{D}^{-1}\mathbf{W}$ is row-stochastic,
$\lambda_1 = 1$ and $\psi_1$ is the diagonal of $\mathbf{D}^{1/2}$.
%
The next few eigenvectors can be argued to describe the geometry of the underlying manifold \cite{coifman2005geometric}.
%
However, some eigenvectors can be higher harmonics of the same principal direction along the data manifold.
This is analogous to how the eigenfunctions $\cos x$ and $\cos 2x$ of the usual Laplacian in one spatial dimension
and with no flux boundary conditions are one-to-one with the values of $x$ for $0 \le x \le 1$;
one must check for correlations between the eigenvectors before selecting those that describe the underlying manifold geometry.
%
The above steps to construct the nonlinear intrinsic variables are summarized in Algorithm \ref{algo}.

@@

\subsection{Limiting Diffusion Operator} \label{sec:diff_limit}
Assume that the dynamics of the underlying variables in \eqref{eq:dynamics} can be rewritten as
\begin{equation} \label{eq:Langevin}
	d\mathbf{x}(t) = - \nabla U(\mathbf{x}(t)) dt + d \mathbf{w}(t),
\end{equation}
where the drift term is expressed as a gradient of a potential field $U$ and $\dot{\mathbf{w}}(t) \in \mathbb{R}^d$ is a vector consisting of independent white noises.
%
Nadler {\em et al.} \cite{nadler2006diffusion} showed that in the limit $\varepsilon \rightarrow 0$ and $T \rightarrow \infty$ the kernel $\mathbf{W}^{\mathrm{norm}}$ converges to the continuous backward Fokker-Planck operator
\begin{equation}
	\mathcal{L} = \Delta - \nabla U \cdot \nabla.
	\label{eq:limiting_operator}
\end{equation}
%where $g:\mathcal{M}_x \rightarrow \mathbb{R}$ is any real function on the domain of the underlying variables.
%Since we assume independent underlying variables, the potential is additive \cite{singer2008non}
%\begin{equation}
%	U(\mathbf{x}(t)) = \sum \limits _{i=1}^d U_i(x_i(t)).
%\end{equation}
We remark that, while in diffusion maps \cite{coifman2005geometric}, the limiting operator is the Laplacian on the manifold on which the observations lie,
here the Laplacian operator on the data that is based on the Mahalanobis distance converges to a limiting Laplacian operator on the
manifold of the intrinsic variables $\mathbf{x}$.

The convergence to the limiting operator implies that the eigenvectors of $\mathbf{W}^{\mathrm{norm}}$, which are used to represent the NIV, approximate the eigenfunctions of the limiting Fokker-Planck operator on the NIV manifold.
%
As such, according to \eqref{eq:limiting_operator}, they depend on the potential field.
%
Thus, the constructed embedding is consistent as long as the subsets of observations $\mathbf{Y}(t)$ are
adequate samples of the same (equilibrium) density.
%
For example, when looking at multiple data sets from a molecular simulation, the embeddings will be consistent provided each simulation sufficiently samples the same underling free energy surface.

In order to merge observations from different regions,
the fundamental question is, to what extent the respective sampling densities affect the coordinates we obtain.
%
In principle, we can decouple diffusion dynamics (and thus, equilibrium densities)
from geometry by normalizing the diffusion kernel with the sample density.
%
This has been explored for diffusion maps independently of NIV \cite{coifman2005geometric}.
%
When the normalized kernel is defined as 
\begin{equation} \label{eq:norm_kernel}
\widetilde{\mathbf{W}}^{\mathrm{norm}} = \mathbf{D}^{-1}\mathbf{W}\mathbf{D}^{-1},
\end{equation}
the limiting diffusion operator is the Laplace-Beltrami operator \cite{nadler2006diffusion}, given by $\widetilde{\mathcal{L}} = \Delta $.
%
In other words, this diffusion captures only the geometry of the observations and is invariant to the potential field (or, alternatively,
to the details of the sampling density).
%
%Therefore, this normalization separates the geometry of the underlying manifold of intrinsic variables from the statistics of the observations.

Since local coordinates are based on the noise term in \eqref{eq:Langevin}, sets of measurements from different regions
will essentially detect the same system of coordinates (the same variable ``axes"), yet different sampling densities (arising, for example,
from the drift term in \eqref{eq:Langevin}) will result in different {\em scalings} of these axes.
%
Applying the Laplace-Beltrami normalization in the NIV context will thus eliminate sampling density effects.
%
An alternative approach would find the scaling transformations that optimally ``register"
data sampled from common regions (in the spirit of histogram reweighting \cite{ferrenberg1988new}).

@@

A na\"{\i}ve way to extend $\mathbf{f}$ to a new data point $\Psi(\tilde{i})$ is find the point's nearest neighbors in NIV space and average their function values.
%
A different, point-wise adaptive approach is described by Buchman {\em et al.} \cite{buchman2011high}:
high-dimensional hurricane tracks were estimated from low dimensional embedding coordinates using a weighted average of the points close to $\Psi(\tilde{i})$ in the embedded space.
%
However, this point-wise adaptation requires setting the nearest neighborhood radius parameter for every point.
%
The LP algorithm finds the appropriate nearest neighborhood radius for each new point $\Psi(\tilde{i})$.
%
This radius will be large in smooth regions of the function, and small in regions in which $\mathbf{f}$ contains higher frequency components.
%
The LP approximation of a new, high dimensional point $\mathbf{Y}(\tilde{i})$ is calculated by a weighted average of the function values that belong to the neighboring points.
%
The weights are based on the pairwise distances in the intrinsic, low-dimensional space.
%
In practice, a set of smoothing operators $P_0, P_1, \ldots, P_l$, with
\begin{equation} \label{eq:LP_multi_scale_app}
P_l \propto e^{-d^2(\Psi(i),\Psi(j)) / 2^{-l} \sigma_0},
\end{equation}
are constructed and later extended to create the multiscale approximations as defined in Eq. \ref{eq:LP_multi_scale}.
%
The LP algorithm for the inverse mapping is summarized in Algorithm \ref{algo_LP}.

@@

\item
Use the smoothing operators to obtain a multiscale representation (see Eq. \ref{eq:LP_multi_scale_app}) of $\mathbf{f}:\Psi(i) \rightarrow Y(i)$.

@@

%\item
%The value $Y(\tilde{i})$ is approximated by $\mathbf{f}_l(\Psi(\tilde{i}))$.

@@

We generate 3000 initial conditions $\mathbf{Y}_0(1), \dots, \mathbf{Y}_0(3000) \in \mathbb{R}^6$ uniformly at random from the region of state space where all concentrations are non-negative.
%
We evolve each point $\mathbf{Y}_0(t)$ forward for 10 time units using the SSA to obtain a point $\mathbf{Y}(t) \in \mathbb{R}^6$;
according to the time scales calculated from the linearized ODEs, 10 time units is sufficiently long for the initial points in the ODE system to converge to the two-dimensional manifold,
but not long enough for the points to converge to a one-dimensional curve or to the final steady state (see Appendix \ref{app:rxn} for more details).
%
In our stochastic simulations, the initial points appear to converge to an approximate two-dimensional manifold
(in expected value, see Figure \ref{fig:rxn_manifolds}).
%
We consider $\mathcal{Y} = \{ \mathbf{Y}(t): t=1, \dots, 3000 \}$ to be representative points ``on" this apparent two-dimensional manifold.
%
From each manifold point $\mathbf{y} \in \mathcal{Y}$, we run 20 short simulation ``bursts'', each for 0.2 time units.
%
We denote the endpoints from the short simulations as $\mathcal{Y}^{burst}(\mathbf{y})$.

@@

To determine whether a sufficient amount of data is available to obtain an accurate embedding, we compute the NIV embeddings for different numbers of data points 
(arising from different numbers of trajectories initialized randomly in the full concentration space).
%
Figure \ref{fig:rxn_convergence} shows the two-dimensional NIV embeddings and corresponding eigenvalue spectra of the kernel for the Gillespie SSA example computed from 10, 100, 500, and 1500 data points. 
%
We observe that the eigenvalue spectra appear converged for 100 data points and above. 
%
This implies that a sufficient amount of data is available merely to discover the intrinsic variables. 
%
However, we observe that 100 points are insufficient to construct a self-consistent embedding {\em in the two variables discovered}. 
%
In order to account for the different scaling of the NIV coordinate axes, we require at least 500 data points. 
%
This last convergence is exemplified in Figure \ref{fig:rxn_convergence} (c) and (d).
%
The convergence of the spectrum of the kernel and the ultimate self-consistency of the NIV embedding is our empirical indicator that a sufficient amount of data is available. 
\begin{figure*}[ht]
    \includegraphics[width=6in]{fig4b}
    \caption{Two-dimensional NIV embeddings (top) and corresponding eigenvalue spectra (bottom) calculated from subsets of $\mathcal{Y}_1$ containing (a) 10 data points, (b) 100 data points, (c) 500 data points, and (d) 1500 data points. One can see that the embedding and spectrum in (a) is inconsistent with the other embeddings. }
    \label{fig:rxn_convergence}
\end{figure*}

@@

We first compare NIV with direct diffusion maps \cite{coifman2005geometric}, an established nonlinear dimensionality reduction technique.
%
We consider 10,000 data points from our simulation $\mathbf{Y}(1), \dots, \mathbf{Y}(10000) \in \mathbb{R}^{30}$; every 100 data points comes from a continuous simulation trajectory.
%
We construct two data sets:
%
$\mathcal{Y}_{even} = \{\mathbf{Y}(t) \text{ restricted to atoms 2, 4, 6, 8, and 10}: t=1, \dots, 10000 \}$,
and $\mathcal{Y}_{odd} = \{\mathbf{Y}(t) \text{ restricted to atoms 5, 7, and 9}: t=1, \dots, 10000 \}$ (see Figure \ref{fig:ala_fes} for the atom indexing).
%
We then compute the NIV and diffusion maps embeddings for $\mathcal{Y}_{even}$ and $\mathcal{Y}_{odd}$;
for NIV, we compute the covariances as in \eqref{eq:cov}, with $L=10$.
%at the ends of each simulation trajectory, we take fewer points before or following the point of interest to estimate the covariance
%(as neighboring points in our data set that do not come from the
%same trajectories should not contribute to the estimate of the local covariance).

@@

%\begin{table}
%\begin{tabular}{| r || c | c |}
%  \hline
%  & Coordinate 1 & Coordinate 2 \\
%  \hline
%  \hline
%  NLICA & 0.62 & 0.84 \\
%  \hline
%  DM & 0.54 & 0.60 \\
%  \hline
%\end{tabular}
%\caption{Correlation between the embedding coordinates from NLICA and DMAPS using the atoms 5, 7, and 9 and using the atoms 2, 4, 6, 8, and 10.}
%\end{table}

We then use NIV together with LP to predict the conformation of Ala2 when we only observe some of the atoms.
%
We have 20000 data points $\mathbf{Y}(1), \dots, \mathbf{Y}(20000)$, where every 100 data points come from one continuous simulation trajectory.
%
Our first data set (which will serve as our training data for LP), $\mathcal{Y}_{all}$,
consists of the first 10000 data points ($\mathcal{Y}_{all} = \{\mathbf{Y}(t): t=1, \dots, 10000\}$).
%
Our second data set (which will serve as our test data), $\mathcal{Y}_{odd}$, consists of the last 12000 data points restricted to {\em only} the odd atoms
($\mathcal{Y}_{odd} = \{ \mathbf{Y}(t) \text{ restricted to the odd atoms}: t = 8001, \dots, 20000\}$).
%
We compute the covariances as in \eqref{eq:cov} with $L=15$.
%at the ends of each simulation trajectory, we take fewer points before or following the point of interest to estimate the covariance.
%
We compute the NIV embedding for the training data $\mathcal{Y}_{all}$ and the test data $\mathcal{Y}_{odd}$; we then use LP interpolation from the training data to predict the location of all the atoms for each point in the test data.

@@

%\begin{table}
%\begin{tabular}{| r || c | c | c |}
%  \hline
%  & Coordinate 1 & Coordinate 2 & Coordinate 3 \\
%  \hline
%  \hline
%  NLICA & 0.97 & 0.72 & 0.85 \\
%  \hline
%  DM & 0.95 & 0.76 & 0.74 \\
%  \hline
%\end{tabular}
%\caption{Correlation between the embedding coordinates from NLICA and DMAPS using all and just odd coordinates.}
%\end{table}

@@

Figure \ref{fig:ala_recon} shows the reconstructed position from partial observation versus true position for certain selected atoms.
%
The strong correlation between the true and reconstructed positions is easier to appreciate for
atoms that move substantially within the data set (such as atoms 1 and 3).
%
Figure \ref{fig:ala_molecules} shows molecular structures for the true and reconstructed configurations for selected data points;
there is qualitative agreement between the true and reconstructed configurations.
%
The discrepancies between the true and reconstructed positions of atoms 9 and 10 (see Figure \ref{fig:ala_fes} for atom indexing) for some of the molecular structures is not surprising.
%
From Figure \ref{fig:ala_embed}, one can see that the first three NIV describe the flipping of atoms 1 and 3 and the dihedral angles $\phi$ and $\psi$.
%
However, atoms 9 and 10 do not participate in any of these physical quantities, and so there is little information about their positions contained in the first three NIV.
%
We suspect that using more NIV would result in a better description of atoms 9 and 10 and lead to more accurrate structure reconstructions.

For a brief comparison of LP over other reconstruction techniques, we also reconstruct configurations from the
NIV components using simple nearest-neighbor interpolation.
%
The average reconstruction error, scaled by the average bond length within the molecule, is shown in Figure \ref{fig:ala_mse};
LP arguably outperforms simple nearest neighbor search for all of the atoms.
%
To determine why LP outperforms nearest neighbor search, we look at the errors at different scales in the LP algorithm.
%
The example described in Figure \ref{fig:LP_ex} demonstrates the appropriateness of the LP algorithm to signals with an intrinsic multiscale structure. 
%
We performed a similar analysis for the data from the alanine dipeptide example, and showed that the signal exhibits a similar error structure, with large errors in some regions of NIV space and small errors in other regions; these results are shown in Figure \ref{fig:LapPyr_ala2_errors}. 
%
We therefore concluded that the alanine dipeptide data does contain a measure of multiscale behavior, and, therefore, that LP is an appropriate algorithm for reconstruction.

@@

\begin{figure*}[ht]
    \centering
    \includegraphics[width=3in]{fig10b}
    \caption{The error in the reconstructed position of the z-coordinate of atom 10 as a function of $\psi_1$. One can see that the error is larger when $\psi_1$ is large, demonstrating that our data is indeed multiscale (similar to Figure \ref{fig:LP_ex} and that Laplacian Pyramids is an appropriate interpolation algorithm to use.}
    \label{fig:LapPyr_ala2_errors}
\end{figure*}

@@

%In addition, we have stated that the measurement function $\mathbf{f}$ has to be ``rich enough'' in order to recover the variables of the underlying stochastic process.
%However, determining whether $\mathbf{f}$ meets such criteria is a non-trivial task.
%For many problems, one would like to say what types of measurement functions $\mathbf{f}$ are rich enough,
%as knowing which functions retain sufficient detail of the underlying stochastic process can aid us in building accurate reduced models.
%
%
Other issues, such as accurately estimating the covariance matrices required for the computation of the Mahalanobis distance, are also of current research interest.
%
It is clearly necessary to link this type of calculation with modern estimation techniques for (multiscale) diffusions \cite{ait2002maximum, ait2003effects, ait2008closed}
to test the appropriateness of the window sampling lengths selected; this will determine the accuracy of the noise covariance estimation by eliminating
the bias due to drift variations.
%
We are confident that the exploration of these open questions will enable the use of our methodology in many  interesting applications,
such as merging data from molecular simulations at different levels of granularity, or merging simulation data with experimental observations.

@@

\bibliography{../../references/references}
%\bibliography{references}
